%
% Copyright (C) 2025 Dimitri Theulings <rector@becclesparish.org.uk>
% ------------------------------------------------------------------
%
% This file may be distributed and/or modified under the
% conditions of the LaTeX Project Public License, either version 1.3c
% of this license or (at your option) any later version.
% The latest version of this license is in:
%
%    http://www.latex-project.org/lppl.txt
%
% and version 1.3c or later is part of all distributions of LaTeX
% version 2008-05-04 or later.
%
% Common Worship: Services and Prayers for the Church of England,
% material from which is included here, is copyright (c)
% The Archbishops' Council, 2000, 2005, Church House Publishing.
%

\documentclass[12pt]{report}
\usepackage[a4paper,margin=2cm]{geometry}
\usepackage[font,titles]{liturgy-cw}
\usepackage{verse}
\usepackage[
	reference/font=\itshape,
	leftmargin=15mm,
	rightmargin=5mm,
	aboveskip=.5ex,
	belowskip=.5ex]{scripture}

\begin{document}
	\chapter*{An Order for Night Prayer (Compline)}
	\section*{Preparation}

	\minister{The Lord almighty grant us a quiet night and a perfect end.}
	\amen

	\minister{Our help is in the name of the Lord}
	\all{who made heaven and earth.}

	\vspace{1ex}

	\rubric{A period of silence for reflection on the past day may follow.}

	\vspace{1ex}

	\rubric{The following or other suitable words of penitence may be used}
	\all{Most merciful God,\\
		we confess to you,\\
		before the whole company of heaven and one another,\\
		that we have sinned in thought, word and deed\\
		and in what we have failed to do.\\
		Forgive us our sins,\\
		heal us by your Spirit\\
		and raise us to new life in Christ. \amen*}

	\vspace{1ex}

	\minister{O God, make speed to save us.}
	\all{O Lord, make haste to help us.}

	\vspace{1ex}

	\glorybe

	\vspace*{1ex}

	\rubric{The following or another suitable hymn may be sung}

	\setlength{\versewidth}{\dimexpr\linewidth-30mm}
	\begin{verse}[\versewidth]
		Before the ending of the day,\\
		Creator of the world, we pray\\
		That you, with steadfast love, would keep\\
		Your watch around us while we sleep.

		From evil dreams defend our sight,\\
		From fears and terrors of the night;\\
		Tread underfoot our deadly foe\\
		That we no sinful thought may know.

		O Father, that we ask be done\\
		Through Jesus Christ, your only Son;\\
		And Holy Spirit, by whose breath\\
		Our souls are raised to life from death.
	\end{verse}

	\newpage

	\section*{The Word of God}

	\subsection*{Psalmody}

	\rubric{One or more of the following psalms may be used.}

	\subsubsection*{Psalm 4}

	\begin{verse}[\versewidth]
		\flagverse{1}Answer me when I call, O God of my righteousness; \Pause\\
		you set me at liberty when I was in trouble;\\
		\vin have mercy on me and hear my prayer.

		\flagverse{2}How long will you nobles dishonour my glory; \Pause\\
		how long will you love vain things and seek after falsehood?

		\flagverse{3}But know that the Lord has shown me his marvellous kindness; \Pause\\
		when I call upon the Lord, he will hear me.

		\flagverse{4}Stand in awe, and sin not; \Pause\\
		commune with your own heart upon your bed, and be still.

		\flagverse{5}Offer the sacrifices of righteousness \Pause\\
		and put your trust in the Lord.

		\flagverse{6}There are many that say, `Who will show us any good?' \Pause\\
		Lord, lift up the light of your countenance upon us.

		\flagverse{7}You have put gladness in my heart, \Pause\\
		more than when their corn and wine and oil increase.

		\flagverse{8}In peace I will lie down and sleep, \Pause\\
		for it is you Lord, only, who make me dwell in safety.
	\end{verse}

	\subsubsection*{Psalm 91}

	\begin{verse}[\versewidth]
		\flagverse{1}Whoever dwells in the shelter of the Most High \Pause\\
		and abides under the shadow of the Almighty,

		\flagverse{2}Shall say to the Lord, ‘My refuge and my stronghold, \Pause\\
		my God, in whom I put my trust.’

		\flagverse{3}For he shall deliver you from the snare of the fowler \Pause\\
		and from the deadly pestilence.

		\flagverse{4}He shall cover you with his wings\\
		\vin and you shall be safe under his feathers; \Pause\\
		his faithfulness shall be your shield and buckler.

		\flagverse{5}You shall not be afraid of any terror by night, \Pause\\
		nor of the arrow that flies by day;

		\flagverse{6}Of the pestilence that stalks in darkness, \Pause\\
		nor of the sickness that destroys at noonday.

		\flagverse{7}Though a thousand fall at your side\\
		\vin and ten thousand at your right hand, \Pause\\
		yet it shall not come near you.

		\flagverse{8}Your eyes have only to behold \Pause\\
		to see the reward of the wicked.

		\flagverse{9}Because you have made the Lord your refuge \Pause\\
		and the Most High your stronghold,

		\flagverse{10}There shall no evil happen to you, \Pause\\
		neither shall any plague come near your tent.

		\flagverse{11}For he shall give his angels charge over you, \Pause\\
		to keep you in all your ways.

		\flagverse{12}They shall bear you in their hands, \Pause\\
		lest you dash your foot against a stone.

		\flagverse{13}You shall tread upon the lion and adder; \Pause\\
		the young lion and the serpent you shall trample underfoot.

		\flagverse{14}Because they have set their love upon me,\\
		\vin therefore will I deliver them; \Pause\\
		I will lift them up, because they know my name.

		\flagverse{15}They will call upon me and I will answer them; \Pause\\
		I am with them in trouble,\\
		\vin I will deliver them and bring them to honour.

		\flagverse{16}With long life will I satisfy them \Pause\\
		and show them my salvation.
	\end{verse}

	\subsubsection*{Psalm 134}

	\begin{verse}[\versewidth]
		\flagverse{1}Come, bless the Lord, all you servants of the Lord, \Pause\\
		you that by night stand in the house of the Lord.

		\flagverse{2}Lift up your hands towards the sanctuary \Pause\\
		and bless the Lord.

		\flagverse{3}The Lord who made heaven and earth \Pause\\
		give you blessing out of Zion.
	\end{verse}

	\rubric{At the end of the psalmody, the following is said or sung}

	\glorybe

	\subsection*{Scripture Reading}

	\rubric{One of the following short lessons or another suitable passage is read}

	\begin{scripture}[Jeremiah 14.9]
		You, O Lord, are in the midst of us and we are called by your name; leave us not, O Lord our God.
	\end{scripture}

	\alternative

	\begin{scripture}[1 Peter 5.8, 9]
		Be sober, be vigilant, because your adversary the devil is prowling round like a roaring lion, seeking for someone to devour. Resist him, strong in the faith.
	\end{scripture}

	\newpage

	\alternative

	\begin{scripture}[Revelation 22.4, 5]
		The servants of the Lamb shall see the face of God, whose name will be on their foreheads. There will be no more night: they will not need the light of a lamp or the light of the sun, for God will be their light, and they will reign for ever and ever.
	\end{scripture}

	\vspace*{2ex}

	\rubric{The following responsory may be said}
	\minister{Into your hands, O Lord, I commend my spirit.}
	\all{Into your hands, O Lord, I commend my spirit.}

	\minister{For you have redeemed me, Lord God of truth.}
	\all{I commend my spirit.}

	\minister{Glory to the Father and to the Son\\
		and to the Holy Spirit.}
	\all{Into your hands, O Lord, I commend my spirit.}

	\minister{Keep me as the apple of your eye.}
	\all{Hide me under the shadow of your wings.}

	\subsection*{Gospel Canticle}

	\rubric{The Nunc dimittis (The Song of Simeon) is said or sung}

	\all{Save us, O Lord, while waking,\\
		and guard us while sleeping,\\
		that awake we may watch with Christ\\
		and asleep may rest in peace.}

	\begin{verse}[\versewidth]
		\flagverse{1}Now, Lord, you let your servant go in peace: \Pause\\
		your word has been fulfilled.

		\flagverse{2}My own eyes have seen the salvation \Pause\\
		which you have prepared in the sight of every people;

		\flagverse{3}A light to reveal you to the nations \Pause\\
		and the glory of your people Israel.\\
		\attrib[50mm]{Luke 2.29-32}
	\end{verse}

	\glorybe

	\vspace{2ex}

	\all{Save us, O Lord, while waking,\\
		and guard us while sleeping,\\
		that awake we may watch with Christ\\
		and asleep may rest in peace.}

	\newpage

	\section*{Prayers}

	\rubric{Intercessions and thanksgivings may be offered here.}

	\subsection*{The Collect}

	\rubric{Silence may be kept.}

	\minister{Visit this place, O Lord, we pray,\\
		and drive far from it the snares of the enemy;\\
		may your holy angels dwell with us and guard us in peace,\\
		and may your blessing be always upon us;\\
		through Jesus Christ our Lord.}
	\amen

	\vspace{1ex}

	\rubric{The Lord’s Prayer may be said.}

	\section*{The Conclusion}

	\minister{In peace we will lie down and sleep;}
	\all{for you alone, Lord, make us dwell in safety.}

	\minister{Abide with us, Lord Jesus,}
	\all{for the night is at hand and the day is now past.}

	\minister{As the night watch looks for the morning,}
	\all{so do we look for you, O Christ.}

	\minister{[Come with the dawning of the day}
	\all{and make yourself known in the breaking of the bread.\normalfont]}

	\minister{The Lord bless us and watch over us;\\
		the Lord make his face shine upon us and be gracious to us;\\
		the Lord look kindly on us and give us peace.}
	\amen

	\vspace*{\fill}

	\begin{center}
		\footnotesize Common Worship: \textit{Services and Prayers for the Church of England}, material from which is included here, \\ is copyright \textcopyright\ The Archbishops' Council, 2000, 2005, Church House Publishing.
	\end{center}


\end{document}